Demystifying the Combination of Dynamic Slicing and \\ Spectrum-based Fault Localization

### 2019-02-25 paper/ijcai/paper.tex
pass on RQ3 (halfway through)

As to quantify complexity of fault localization on this benchmark, we
counted the number of cases of ``fault of omission'', a popular metric
in fault localization research for that
purpose~\cite{Pearson:2017:EIF:3097368.3097441}.  A fault of omission
(\foo) occurs when the program is buggy because some code is missing,
i.e., a fix to the bug iss created by adding code as opposed to
modifying or changing code. Intuitively, the ability of techniques to
produce accurate diagnostic reports in those cases might be poor as
techniques can only extract information about existing statements,
which are not problematic. This dataset has a total of $59.7\%$ of
\foo s, making it challenging for techniques to perform
well.

@@

A necessary but insufficient condition for improvement of fault
localization with \comb{} is that dynamic slicing eliminates a high
amount of code in the best ranked files. Table \ref{tab:red} shows, as
percentages, the reduction of the top file sizes
obtained with dynamic slicing for different values of
$k$. Overall, results indicate
that dynamic slicing substantially reduces the size of highly-ranked
files.

@@

A fundamental issue of purely dynamic slicing techniques is the risk
of discarding faulty statements (see Section~\ref{sec:slicing}). This
research question evaluates the practical impact of this conceptual
issue. Table~\ref{tab:ds-cases-captured} shows the number of cases
dynamic slicing captures at least one of the faulty statements among
the sliced files for different values of $k$. These numbers indicate
that in general \ds{} finds at least one faulty statement in the top
$5$ of the highest-ranked faulty classes ($87.3\%$ on average) which
is slightly improved to $91.2\%$ if we consider
the top $10$ of the highest-ranked faulty classes.

@@

		\bottomrule
	\end{tabular}
	\caption {\ds{} performance on capturing faulty statements, as
     percentages. Higher is better.}
   \label{tab:ds-cases-captured}
\vspace{-0.2cm}

@@

To analyze the impact of dynamic slicing in context, we also compared
\combpar{k} and \sfl{}$^{k}$ considering the number of cases where the
fault is captured by the technique. Note that the reason \sfl{}$^{k}$
misses the fault is different compared to \combpar{k}. \sfl{}$^{k}$
captures the fault if the fault is included within the top $k$ files;
it misses the fault otherwise. Table~\ref{table:fsws} shows results,
indicating that \combpar{k}, on average, captures more faults compared
to \sfl{}$^{k}$. In only two cases \sfl{}$^{k}$ outperformed
\combpar{k}. These results show that the majority of faulty statements
are found on the $5$ highest ranked faulty classes and that
\combpar{10} is the variant of our approach that performs better
missing only $23$ out of the \numFaults\ ($\sim$8.8\%) faults.

@@

\begin{table}[h]
	\vspace{-0.2cm}
	\small
	\setlength{\tabcolsep}{3pt}
	\centering
\begin{tabular}{l|rr|rr}
		 \toprule
       & \multicolumn{2}{c|}{$k=5$} & \multicolumn{2}{c}{$k=10$} \\
		 Project        &   \sfl{}   & \comb &   \sfl{}   & \comb\\
		\midrule
		 \lang{}         &   84.6   & 100.0 & 84.6 & 96.9  \\
		\cmath{}           & 81.7 & 89.4 & 85.6 & 95.2 \\
		\chart{}           & {\cellcolor{Gray} 84.6} & {\cellcolor{Gray}76.9} & {\cellcolor{Gray}92.3} & {\cellcolor{Gray}84.6}\\
		\jtime{}           & 77.8 & 81.5 & 81.5 & 85.2\\
		\mockito{}           & 63.2 & 71.1 & 71.1 & 78.9\\   \midrule
		Total         & 79.6 & 87.3 & 83.5 & 91.2\\
		\bottomrule
	\end{tabular}
  \caption {Number of faults where at least one of the faulty-statements is in the \sfl{}$^{k}$ and \combpar{k} report}
  \label{table:fsws}
\vspace{-0.4cm}
\end{table}

@@

In \textbf{RQ3}, we evaluate the effective impact of combining \ds{}
with \sfl{} to improve fault localization. The first experiment we ran 
to assess impact consists of measuring the \emph{difference of diagnosis 
cost}. This metric has been previously used in other studies~\cite{7390282,ang-perez-van-deursen-rui-2017,Pearson:2017:EIF:3097368.3097441,Perez:2018:LQR:3304889.3304927}. 
More specifically, we computed
$\Delta{}C = C(\textrm{SFL$^{k}$}) - C(\textrm{\combpar{k}})$, where
$C$ denotes the diagnosis cost and is obtained by computing the mean
position in the ranking of all buggy statements. $\Delta C >0$ means
that the position of the faulty statement has decreased with
\combpar{k}, \ie{}, the combination had a positive impact. $\Delta C
<0$ means that \combpar{k} performs worse compared to its
baseline. That could happen if dynamic slicing misses the faulty
statement---the statements not included in the slice will have their
suspiciousness scores reduced and ranks potentially reduced. $\Delta
C=0$ means that the faulty statement remained in the same position.
It is worth noting that we assume a perfect bug understanding, i.e.,
if the developer sees the bug in the ranking, he is able to precisely
determine if it is the faulty one.

Figure \ref{fig:diagnosis} summarizes performance of \combpar{k}
relative to \sfl{}$^{k}$ using histograms over $\Delta{}C$ for
different values of $k$. Note that (1) the number of cases where
\sfl{}$^{k}$ outperforms \combpar{k} is very small compared and (2)
the number of cases \combpar{k} outperforms \sfl{}$^{k}$ is
significant and, as we increase the value of $k$, it increases (from
$97$ to $117$ faults). The faults where the cost of diagnosis is
negative are cases that (1) the faulty statements decreased in the ranking
list due to being a fault of omission, (2) the test has design limitations
which \ds~is not capable of handling, or (3) the buggy-class in not in
the top $k$ of the highest-ranked faulty classes under analysis.

@@

>>>>

%% a popular metric in fault
%% localization research for that
%% purpose.  A fault of omission
%% (\foo) occurs when the program is buggy because some code is missing,
%% i.e., a fix to the bug is created by adding code as opposed to
%% modifying or changing code.

As to quantify complexity of fault localization on this benchmark, we
measured the frequency of ``faults of omission'' (\foo), which is a
fault whose fix is materialized only with the addition of new
code~\cite{Pearson:2017:EIF:3097368.3097441}. Intuitively, the ability
of techniques to produce accurate diagnostic reports in those cases
might be poor as existing statements are not problematic. We found
that this dataset has a total of $59.7\%$ of \foo s, making it
challenging for techniques to perform well.

@@

A necessary but insufficient condition for improvement of fault
localization with \comb{} is that dynamic slicing eliminates a high
amount of code from the best ranked files. Table \ref{tab:red} shows,
as percentages, the size reduction of the top files obtained with
dynamic slicing for different values of $k$. Overall, results indicate
that dynamic slicing substantially reduces the size of highly-ranked
files. \Mar{+ coverage?}

@@

A fundamental issue of purely dynamic slicing techniques is the risk
of discarding faulty statements (see Section~\ref{sec:slicing}). This
research question evaluates the practical impact of this conceptual
issue. Table~\ref{tab:ds-cases-captured} shows the number of cases
dynamic slicing captures at least one of the faulty statements among
the sliced files for different values of $k$. These numbers indicate
that in $87.3\%$ of the cases, on average, dynamic slicing finds at
least one faulty statement in the top $5$ of the highest-ranked faulty
classes. This number is slightly improved to $91.2\%$ if we consider
the top $10$ of the highest-ranked faulty classes.

@@

		\bottomrule
	\end{tabular}
	\caption {Dynamic Slicing performance on capturing faulty statements, as
     percentages. Higher is better.}
   \label{tab:ds-cases-captured}
\vspace{-0.2cm}

@@

To analyze the impact of dynamic slicing in context, we also compared
\combpar{k} and \sfl{}$^{k}$ considering the number of cases where the
fault is captured by the technique within a certain bound $k$. Note
that the reason \sfl{}$^{k}$ misses the fault is different compared to
\combpar{k}. \sfl{}$^{k}$ captures the fault if it is included within
the top $k$ files; it misses the fault
otherwise. Table~\ref{table:fsws} shows results, indicating that
\combpar{k}, typically, captures more faults compared to
\sfl{}$^{k}$. In only two cases, highlighted in gray color,
\sfl{}$^{k}$ outperformed \combpar{k}. These results show that the
majority of faulty statements are found on the $5$ highest ranked
faulty classes and that \combpar{10} is the variant of our approach
that performs better missing only $23$ out of the
\numFaults\ ($\sim$8.8\%) faults.

@@

\begin{table}[h]
	\vspace{-0.2cm}
	\small
	\setlength{\tabcolsep}{3pt}
	\centering
\begin{tabular}{l|rr|rr}
		 \toprule
       & \multicolumn{2}{c|}{$k=5$} & \multicolumn{2}{c}{$k=10$} \\
		 Project        &   \sfl{}   & \comb &   \sfl{}   & \comb\\
		\midrule
		 \lang{}         &   84.6   & 100.0 & 84.6 & 96.9  \\
		\cmath{}           & 81.7 & 89.4 & 85.6 & 95.2 \\
		\chart{}           & {\cellcolor{Gray} 84.6} & {\cellcolor{Gray}76.9} & {\cellcolor{Gray}92.3} & {\cellcolor{Gray}84.6}\\
		\jtime{}           & 77.8 & 81.5 & 81.5 & 85.2\\
		\mockito{}           & 63.2 & 71.1 & 71.1 & 78.9\\   \midrule
		Total         & 79.6 & 87.3 & 83.5 & 91.2\\
		\bottomrule
	\end{tabular}
  \caption {Number of faults where at least one of the faulty
    statements appears at the report of the technique. Higher is better.}
  \label{table:fsws}
\vspace{-0.4cm}
\end{table}

@@

In \textbf{RQ3}, we evaluate the impact of combining \ds{} with \sfl{}
to improve fault localization. The first experiment we ran consists of
measuring the \emph{difference of diagnosis cost} between the two
fault localization techniques. This metric has been previously used in
other
studies~\cite{7390282,ang-perez-van-deursen-rui-2017,Pearson:2017:EIF:3097368.3097441,Perez:2018:LQR:3304889.3304927}.
More specifically, we computed $\Delta{}C = C(\textrm{SFL$^{k}$}) -
C(\textrm{\combpar{k}})$, where $C$ denotes the diagnosis cost and is
obtained by computing the mean position in the ranking of all buggy
statements. $\Delta C >0$ means that \combpar{k} outperformed the
baseline, \ie{} the mean position of the faulty statements is smaller
in \combpar{k} compared to that of the baseline. $\Delta C <0$ means
that \combpar{k} performs worse compared to its baseline. That could
happen if dynamic slicing misses the faulty statement. $\Delta C=0$
means that the faulty statement remained in the same position.  It is
worth noting that we assume perfect bug
understanding~\cite{Parnin:2011:ADT:2001420.2001445}, i.e., if the
developer sees the bug in the ranking, he is able to precisely
determine if it is the faulty one.

Figure \ref{fig:diagnosis} summarizes performance of \combpar{k}
relative to \sfl{}$^{k}$. The figure shows a histogram over
$\Delta{}C$ for different values of $k$. Note that (1) the number of
cases where the baseline outperforms \combpar{k} is small (\eg{}, $33$
of the \numFaults\ faults for $k=10$) and that (2) the number of cases
\combpar{k} outperforms the baseline is significative (\eg{}, $117$ of
the \numFaults\ faults for $k=10$). One of the following reasons
justifies the baseline outperforming \combpar{k}:\Mar{following
  sentence is confusing IMO. suggest to clarify or remove. aren't
  points two and three equivalent? as for one, isn't foo an issue both
  for \combpar{k} and \sfl{}$^{k}$?} (1) a fault of omission, (2) the
test has design limitations which \ds~is not capable of handling, or
(3) the buggy-class in not in the top $k$ of the highest-ranked faulty
classes under analysis.

@@

\Mar{-------------------------- estou aqui}
